---
title: "ggsoiltexture"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of ggsoiltexture is to provide a simple ggplot function for the plotting of soil textural data. It is still in development and under review in a journal for publication. Meanwhile, if you use this package please cite this repository. Hope it is useful!

## Acknowledgements 
The code was development based on the ggplot_piper functions written by [Jonh Dorian](https://gist.github.com/johnDorian/5561272) and inspired by the R package [ggtern](https://github.com/nicholasehamilton/ggtern). Thanks for sharing your knowledge. Also, thanks for the X users that give us feedback about functionalities and new ideas.

## Installation

You can install the last version of ggsoiltexture from [GitHub](https://github.com/Saryace/ggsoiltexture). This package will be available in CRAN after publication in a journal. 


```{r install, message = FALSE}
# install.packages("devtools")
# devtools::install_github("Saryace/ggsoiltexture")
```

## Main function: ggsoiltexture

You can plot a simple soil textural triangle directly:

```{r simple_plot, warning=FALSE}
library(tidyverse)
library(ggsoiltexture)
some_data <- data.frame(id = c("Soil A","Soil B","Soil C","Soil D","Soil E"),
                          clay = c(10,20,25,20,50),
                          silt  = c(35,15,45,30,10),
                          sand = c(55,65,30,50,40),
                          om = c(5,15,5,12,7))

ggsoiltexture(some_data)
```

Also, it is posible to remove the grid:

```{r simple_plot-2}
ggsoiltexture(some_data,
              show_grid = FALSE)
```

Because this function is based on ggplot, you can use ggplot2 functionalities and other packages

```{r simple_plot-3}
library(ggrepel)
pub_plot <-
    ggsoiltexture(some_data) +
    geom_point(aes(color = om), size = 6) +
    scale_color_continuous(type = "viridis") +
    labs(color = "Organic\nMatter (%)") +
    geom_label_repel(aes(label = id), box.padding = 0.5) +
    theme(legend.title = element_text(face = "bold"),
          legend.position = "bottom")

pub_plot
```

ggsoiltexture includes some classification systems from USDA, New Zeland, Germany and Swiss.

```{r simple_plot-4}
ggsoiltexture(some_data,
              show_grid = FALSE,
              class = "USDA")
```

The fontsize of the textural class can be changed by the geom_polygon layer:

```{r simple_plot-a}

```


```{r simple_plot-5}
ggsoiltexture(some_data,
              show_grid = FALSE,
              class = "GERMAN")
```

```{r simple_plot-6}
ggsoiltexture(some_data,
              show_grid = FALSE,
              class = "NZ")
```

```{r simple_plot-7}
ggsoiltexture(some_data,
              show_grid = FALSE,
              class = "SWISS")
```
