---
title: "ggsoiltexture"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of ggsoiltexture is to provide a simple ggplot function for the plotting of soil textural data. It is still in development and under review in a journal for publication. Meanwhile, if you use this package please cite this repository. Hope it is useful!

## Acknowledgements 
The code was development based on the ggplot_piper functions written by [Jonh Dorian](https://gist.github.com/johnDorian/5561272) and inspired by the R package [ggtern](https://github.com/nicholasehamilton/ggtern). Thanks for sharing your knowledge. Also, thanks for the X users that give us feedback about functionalities and new ideas.

## Installation

You can install the last version of ggsoiltexture from [GitHub](https://github.com/Saryace/ggsoiltexture). This package will be available in CRAN after publication in a journal. 


```{r install, message = FALSE}
# install.packages("devtools")
# devtools::install_github("Saryace/ggsoiltexture")
```

## Main function: ggsoiltexture

You can plot a simple soil textural triangle directly:

```{r simple_plot, warning=FALSE}
library(tidyverse)
library(ggsoiltexture)
some_data <- tibble(id = c("Soil A","Soil B","Soil C","Soil D","Soil E"),
                          clay = c(10,20,25,20,50),
                          silt  = c(35,15,45,30,10),
                          sand = c(55,65,30,50,40),
                          om = c(5,15,5,12,7))

ggsoiltexture(some_data)
```

Also, it is posible to remove the grid:

```{r simple_plot-2}
ggsoiltexture(some_data,
              show_grid = FALSE)
```

Because this function is based on ggplot, you can use ggplot2 functionalities and other packages

```{r simple_plot-3}
library(ggrepel)
pub_plot <-
    ggsoiltexture(some_data) +
    geom_point(aes(color = om), size = 6) +
    scale_color_continuous(type = "viridis") +
    labs(color = "Organic\nMatter (%)") +
    geom_label_repel(aes(label = id), box.padding = 0.5) +
    theme(legend.title = element_text(face = "bold"),
          legend.position = "bottom")

pub_plot
```

## Adding classification classes

ggsoiltexture includes some classification systems from USDA, New Zealand, Germany and Swiss.

```{r simple_plot-4}
ggsoiltexture(some_data,
              show_grid = FALSE,
              class = "USDA")
```


```{r simple_plot-5}
ggsoiltexture(some_data,
              show_grid = FALSE,
              class = "GERMAN")
```

```{r simple_plot-6}
ggsoiltexture(some_data,
              show_grid = FALSE,
              class = "NZ")
```

```{r simple_plot-7}
ggsoiltexture(some_data,
              show_grid = FALSE,
              class = "SWISS")
```

# Some examples

## Comparing two classification systems. 

The package has the polygon data for each type of classification, so they can be used separately for more advanced plotting. Due to the amount of information, it is recommended to remove all legends and add a description of the figure.

```{r simple_plot-8}

data("nz_polygons")
data("usda_polygons")

ggsoiltexture(tibble(sand = 20,
                     silt = 15,
                     clay = 65),
              show_grid = FALSE,
              class = NULL) +
  geom_polygon(data = nz_polygons, 
               aes(x = x, y = y, fill = label),
               alpha = .5,
               color = "grey90") +
  geom_text(data = nz_polygons %>%
              dplyr::group_by(CODE) %>%
              dplyr::summarise_if(is.numeric, mean, na.rm = TRUE),
            aes(label = CODE, color = CODE),
            show.legend=FALSE,
            size = 3.5) +
  geom_polygon(data = usda_polygons, 
               aes(x = x, y = y, fill = label),
               alpha = 0,
               linetype="dotted",
               color = "grey20") +
  geom_text(data = usda_polygons %>%
              dplyr::group_by(label) %>%
              dplyr::summarise_if(is.numeric, mean, na.rm = TRUE),
            aes(label = label),
            size = 3.5) +
  guides(color = "none", fill = "none") +
  labs(caption = "Dotted lined polygons: USDA classification\nColored polygons: New Zealand Classification")
```

## Using Munsell colors

Munsell colors are widely used for soil classification. Using the library ´munsell´ the color can be transformed to hexcolor and be plotted. 

```{r simple_plot-9}
library(munsell)
library(ggrepel)
munsell_data <- tibble(sand = c(45,60,30,50,70,40,55,65,35,25,80,20),
                       clay = c(20,15,25,28,10,30,22,12,58,35,8,70),
                       silt = 100 - sand - clay,
                       om = c(2.3,1.8,3.2,2.0,1.5,2.6,2.8,1.2,3.0,4.0,1.0,4.2),
                       om_level = ifelse(om > 2, "high", "low"),
                       munsell = c("10YR 3/4","7.5YR 4/6","7.5YR 4/6", "10YR 4/2","5YR 3/6",
                                   "10YR 5/8","7.5YR 5/4","2.5Y 4/2","10YR 6/6",
                                   "5YR 4/8","7.5YR 6/2", "10YR 7/4"))

munsell_data$color_hex <- mnsl(munsell_data$munsell)

plot_munsell <- 
ggsoiltexture(munsell_data) +
  scale_color_identity() +
  geom_point(aes(color = color_hex, size = om, group = om_level)) +
  geom_label_repel(aes(label = munsell), box.padding = 0.5) 

plot_munsell

```

Even it is possible to facet them, due to is based on ggplot :)

```{r}
plot_munsell +
  facet_wrap(vars(om_level))
```

## Cats

Because of cats. This example is just for motivate to the ggsoiltexture's users to explore any idea and try to plot it. 

```{r}
library(ggcats)
library(gganimate)

ggsoiltexture(tibble(sand = c(90, 30, 50, 0, 50),
                     silt = c(5, 30, 30, 33, 10),
                     clay = c(5, 40, 20, 67, 40)),
              class = NULL) +
  geom_cat(aes(cat = sand), cat = "shironeko", size = 6) +
   transition_reveal(sand)

```

# How to contribute

We welcome any issues and pull requests.

Please report any bugs you find by submitting an issue on github.com/Saryace/ggsoiltexture/issues.

If you wish to contribute, please submit a pull request following the guidelines stated in CONTRIBUTING.md.

